---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Retrieving Data from a Custom Web API  

<!-- badges: start -->
[![R-CMD-check](https://github.com/coatless-r-n-d/r-web-api/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coatless-r-n-d/r-web-api/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->
  
The `cumtd` _R_ package provides an example of implementing and using a 
Web API. In particular, the package is a case study in retrieving data from the
[Champaign-Urbana Mass Transit District ('CUMTD')](https://developer.cumtd.com/)
Web API, which provides realtime information for buses. Writing a custom wrapper
to the API was necessary since there is not a realtime [General Transit Feed Specification (GTFS)](http://code.google.com/transit/spec/transit_feed_specification.html)
available via [GTFS Realtime](https://developers.google.com/transit/gtfs-realtime/).
Thus, for those wanting to know if their bus will arrive on time, one must look
elsewhere.

As a follow-up, the package is used in 
[`r-shinydashboard`](https://github.com/r-pkg-examples/r-shinydashboard) example
to illustrate how a real-time API can be used with 
[`shinydashboard`](https://rstudio.github.io/shinydashboard/index.html).

### Installation

To install the `cumtd` package, use:

```r
if(!requireNamespace("remotes")) install.packages("remotes")
remotes::install_github("coatless-r-n-d/r-web-api")
```

To access its contents, load it into _R_ with:

```r
library("cumtd")
```

### Endpoints

We have support presently for:

```{r, echo = FALSE}
# Must be in the environment... 
library("cumtd")
# Dynamically retrieve package functions
funcs = ls("package:cumtd")
funcs = matrix(funcs, nrow = length(funcs))
colnames(funcs) = c("Implemented Functions")
knitr::kable(funcs)
# Shows parameter call
# lsf.str("package:ghapi3")
```

End points to be added (PRs appreciated!):

```{r, echo = FALSE}
ep = data.frame(Endpoints = c("GetCalendarDatesByDate", "GetCalendarDatesByService",
                         "GetNews", 
                         "GetReroutes", "GetReroutesByRoute",
                         "GetRoute", "GetRoutesByStop",
                         "GetStop", "GetStopsByLatLon", "GetStopsBySearch",
                         "GetStopTimesByTrip", "GetStopTimesByStop",
                         "GetPlannedTripsByLatLon", "GetPlannedTripsByStops",
                         "GetTrip", "GetTripsByBlock", "GetTripsByRoute",
                         "GetVehicle", "GetVehiclesByRoute",
                         "GetLastFeedUpdate"))

ep$`End Point Documentation` = paste0("[", ep$Endpoints,"]",
                "(https://developer.cumtd.com/documentation/v2.2/method/", tolower(ep$Endpoints), ")")

knitr::kable(ep[, -1, drop = FALSE])
```


### Implementation Details

When trying to import data into _R_ from a Web API, the preferred approach
is to use [`httr`](https://cran.r-project.org/package=httr) or [`curl`](https://cran.r-project.org/package=curl). Using either of these
packages, one can form HTTP requests to send to the Web API and receive
HTTP responses from it that contain the data. Inside the response, there is
a body portion that contains the desired data and needs to be converted into an
_R_ data structure.


### License

GPL (\>= 2)
